---
title: "`Rvision` - a basic (for now) computer vision library for R"
author: "Simon Garnier - <a href='https://twitter.com/sjmgarnier'>@sjmgarnier</a>"
date: "`r Sys.Date()`"
output:
  html_document:
    css: custom.css
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    number_sections: yes
    theme: cosmo
    highlight: pygments
    toc: yes
  pdf_document:
    toc: yes
vignette: |
  %\VignetteIndexEntry{Rvision computer vision library}
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
---

---

# Package installation

## Requirements

Before installing `Rvision`, you will need to install the latest version of the 
`devtools` package in `R`.You can install `devtools` as follows: 

```{r, eval = FALSE}
install.package("devtools")
```

## Installing Rvision

You can install `Rvision` as follows:

```{r, eval = FALSE}
devtools::install_github("swarm-lab/Rvision")
```

`Rvision` depends on `ROpenCVLite` to access `OpenCV`'s functionalities. If not
already installed, `ROpenCVLite` will be installed first by the above command 
line. This may take some time as it will download, compile and install `OpenCV` 
for you (compilation time will depend on your computer). I suggest going out for
a cup of tea or coffee while `ROpenCVLite` is installing ;-)

[Back to top]("#")

---

# Image operations

## Read an image file

```{r, message=FALSE}
library(Rvision)

# Locate sample file
samplePath <- system.file("sample_img", "bunny.png", package = "Rvision")

# Read image file
img1 <- image(samplePath)
img1
```

## Display image

```{r, fig.height=8*720/1280}
plot(img1)
```

## Image properties

```{r}
dim(img1)
nrow(img1)
ncol(img1)
nchan(img1)
bitdepth(img1)
colorspace(img1)
```

## Save image to disk

```{r}
write.Image(img1, file = "~/Desktop/img1.png")
```

## Convert color image to grayscale

```{r, fig.height=8*720/1280}
img2 <- changeColorSpace(img1, "GRAY")
plot(img2)
```

## Convert grayscale image to black and white

```{r, fig.height=8*720/1280}
img3 <- img2 > 128   # Gray values > 128 are turned to white (255)
                    # Gray values <= 128 are turned to black (0)
plot(img3)
```

[Back to top]("#")

---











