---
title: "`Rvision` - a basic (for now) computer vision library for R"
author: "Simon Garnier - <a href='https://twitter.com/sjmgarnier'>@sjmgarnier</a>"
date: "`r Sys.Date()`"
output:
  html_document:
    css: custom.css
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    number_sections: yes
    theme: cosmo
    highlight: pygments
    toc: yes
  pdf_document:
    toc: yes
vignette: |
  %\VignetteIndexEntry{Vignette Title} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
---

---

# Package installation

## Requirements

Before installing `Rvision`, you will need to install the latest version of the
[`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite) and `devtools` packages
in `R`. 

You can install `devtools` as follows. 

```{r, include=FALSE}
install.packages("devtools")
```

Installation instructions for `ROpenCVLite` can be found here:
[https://github.com/swarm-lab/ROpenCVLite#2-package-installation](https://github.com/swarm-lab/ROpenCVLite#2-package-installation). 
This may take some time as it will download, compile and install `OpenCV` for 
you (compilation speed will depend on your computer).

## Installing Rvision

```{r, include=FALSE}
devtools::install_github("swarm-lab/Rvision")
```

[Back to top]("#")

---

# Image operations

## Read an image file

```{r, message=FALSE}
library(Rvision)

# Locate sample file
samplePath <- paste0(find.package("Rvision"), "/sample_img/bunny.png")

# Read image file
img1 <- image(samplePath)
img1
```

## Display image

```{r, fig.height=8*720/1280}
plot(img1)
```

## Image properties

```{r}
dim(img1)
nrow(img1)
ncol(img1)
nchan(img1)
```

## Save image to disk

```{r}
write.Image(img1, file = "~/Desktop/img1.png")
```

## Convert color image to grayscale

```{r, fig.height=8*720/1280}
img2 <- convert(img1, "gray")
plot(img2)
```

## Convert grayscale image to black and white

```{r, fig.height=8*720/1280}
img3 <- img2 > 75   # Gray values > 75 are turned to white (255)
                    # Gray values <= 75 are turned to black (0)
plot(img3)
```

[Back to top]("#")

---











