---
title: "1 - Installing Rvision"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

__This page describes the installation process of [`Rvision`](https://github.com/swarm-lab/Rvision).__ 

Installing [`Rvision`](https://github.com/swarm-lab/Rvision) should go smoothly 
on all major operating systems (Mac, Windows, "Ubuntu-like" Linux). There are a 
few steps to follow before installing the package itself in order to install all 
the necessary dependencies. 

All installation steps are described below. 

---

+ [1.1 - Before installing ](#before-installing)
    + [1.1.1 - Devtools](#devtools)
    + [1.1.2 - ROpenCVLite](#ropencvlite)
+ [1.2 - Installing Rvision](#installing-rvision)
+ [1.3 - Loading Rvision](#loading-rvision)
  
---

## 1.1 - Before installing

### 1.1.1 - Devtools

Before installing [`Rvision`](https://github.com/swarm-lab/Rvision), you will 
need to install the latest version of the [`devtools`](https://CRAN.R-project.org/package=devtools) 
package. You can install [`devtools`](https://CRAN.R-project.org/package=devtools) 
from [`CRAN`](https://cran.r-project.org) as follows: 

```{r, message=FALSE, eval=FALSE}
if (!require(devtools))
  install.packages("devtools")
```

### 1.1.2 - ROpenCVLite

You will also need to install the [`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite)
package from its [Github repository](https://github.com/swarm-lab/ROpenCVLite). 
Installation instructions for [`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite)
can be found at https://swarm-lab.github.io/ROpenCVLite/articles/install.html.

This step may take some time as it will download, compile and install 
[`OpenCV`](https://opencv.org) for you. I suggest you go out for a cup of tea or 
coffee while [`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite) is being 
installed (compilation time will depend on your computer). 

Note that you can skip this step and [`Rvision`](https://github.com/swarm-lab/Rvision)
will attempt to install it in the next step. However it might fail if your 
system is missing one of the external dependencies required to compile [`OpenCV`](https://opencv.org). 

---

## 1.2 - Installing Rvision

You can install the latest development version of 
[`Rvision`](https://github.com/swarm-lab/Rvision) as follows:

```{r, message=FALSE, eval=FALSE}
devtools::install_github("swarm-lab/Rvision")
```

---

## 1.3 - Loading Rvision

```{r, message=FALSE, eval=FALSE}
library(Rvision)
```
