---
title: "1 - Installing Rvision"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{1 - Installing Rvision}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 1.1 - Before installing

### 1.1.1 - Devtools

Before installing `Rvision`, you will need to install the latest version of the 
[`devtools`](https://CRAN.R-project.org/package=devtools) package. You can 
install `devtools` from [`CRAN`](https://cran.r-project.org) as follows: 

```{r, message=FALSE, eval=FALSE}
if (!require(devtools))
  install.packages("devtools")
```

### 1.1.2 - ROpenCVLite

You will also need to install the [`ROpenCVLite`](https://github.com/swarm-lab/ROpenCVLite)
package. Installation instructions for `ROpenCVLite` can be found at 
https://swarm-lab.github.io/ROpenCVLite/articles/install.html.

This step may take some time as it will download, compile and install 
[`OpenCV`](https://opencv.org) for you. I suggest you go out for a cup of tea or 
coffee while `ROpenCVLite` is being installed (compilation time will depend on 
your computer). 

Note that you can skip this step and `Rvision` will attempt to install it in the 
next step. However it might fail if your system is missing one of the external 
dependencies required by `ROpenCVLite`. 

---

## 1.2 - Installing Rvision

You can install the latest development version of `Rvision` as follows:

```{r, message=FALSE, eval=FALSE}
devtools::install_github("swarm-lab/Rvision")
```

---

## 1.3 - Loading Rvision

```{r, message=FALSE, eval=FALSE}
library(Rvision)
```
