---
title: "2 - Input/output operations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(Rvision)
```

__This page describes how to read and write videos and images with 
[`Rvision`](https://github.com/swarm-lab/Rvision).__ 

---

+ [2.1 - Reading ](#reading)
    + [2.1.1 - Videos](#videos)
    + [2.1.2 - Camera streams](#camera-streams)
    + [2.1.3 - Images](#images)
        + [2.1.3.1 - From files](#from-files)
        + [2.1.3.2 - From videos](#from-videos)
        + [2.1.3.3 - From streams](#from-camera-streams)
+ [2.2 - Writing](#writing)
    + [2.2.1 - Videos](#videos-1)
    + [2.2.2 - Images](#images-1)

---

## 2.1 - Reading

### 2.1.1 - Videos

You can create a `Video` object from a video file as follows: 

```{r, eval=FALSE}
path_to_video <- system.file("sample_vid", "Balloon.mp4", package = "Rvision")
my_video <- video(filename = path_to_video)
```

`Video` objects can be released from memory as follows: 

```{r, eval=FALSE}
release(my_video)
```

Note that `Video` objects are wrappers around [`OpenCV`](http://opencv.org/)
[`VideoCapture`](https://docs.opencv.org/3.4.1/d8/dfe/classcv_1_1VideoCapture.html)
objects. They will not persist between [`R`](https://cran.r-project.org) sessions. 

### 2.1.2 - Camera streams

You can create a `Stream` object from a camera stream as follows: 

```{r, eval=FALSE}
my_stream <- stream(index = 0)
```

The index argument takes an integer number corresponding to the position of the 
camera in the list of video capturing devices available on your computer. 0 
corresponds to the default camera, which is usually the embedded webcam on most 
computers. Note that the order of the list of video capturing devices might 
change after each computer restart. 

`Stream` objects can be released from memory as follows: 

```{r, eval=FALSE}
release(my_stream)
```

Note that `Stream` objects are wrappers around [`OpenCV`](http://opencv.org/)
[`VideoCapture`](https://docs.opencv.org/3.4.1/d8/dfe/classcv_1_1VideoCapture.html)
objects. They will not persist between [`R`](https://cran.r-project.org) sessions. 

### 2.1.3 - Images

There are three ways to create `Image` objects: from files, from `Video` objects,
and from `Stream` objects. 

Note that `Image` objects are wrappers around [`OpenCV`](http://opencv.org/)
[`Mat`](https://docs.opencv.org/3.4.1/d3/d63/classcv_1_1Mat.html) objects. They 
will not persist between [`R`](https://cran.r-project.org) sessions.

#### 2.1.3.1 - From files

You can create an `Image` object from an image file as follows: 

```{r, eval=FALSE}
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")
my_image <- image(filename = path_to_image)
```

#### 2.1.3.2 - From videos

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
path_to_video <- system.file("sample_vid", "Balloon.mp4", package = "Rvision")
my_video <- video(filename = path_to_video)
```

You can create an `Image` object from the next available frame of a `Video` 
object as follows: 

```{r, eval=FALSE}
my_image <- readNext(my_video)
```

You can create an `Image` object from any arbitrary frame of a `Video` object as 
follows (here frame number 100):

```{r, eval=FALSE}
my_image <- readFrame(my_video, 100)
```

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
release(my_video)
```

#### 2.1.3.3 - From camera streams

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
my_stream <- stream(index = 0)
```

You can create an `Image` object from the next available frame of a `Stream` 
object as follows: 

```{r, eval=FALSE}
my_image <- readNext(my_stream)
```

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
release(my_stream)
```

---

## 2.2 - Writing

### 2.2.1 - Videos

Videos can be written to the disk using a `VideoWriter` object. 

When creating a `VideoWriter` object, you need to specify the codec of the video, 
its framerate, its height and its width. For instance, we will here create a 
`VideoWriter` object that will save to a temporary [`mp4`](https://en.wikipedia.org/wiki/MPEG-4_Part_14) 
video file using the [`x264`](https://www.videolan.org/developers/x264.html) 
codec at 30fps. 

```{r, eval=FALSE}
path_to_video <- paste0(tempfile(), ".mp4")
my_writer <- videoWriter(path_to_video, fourcc = "x264", fps = 30, height = 720, width = 1280)
```

Once a `VideoWriter` object has been created, you can write individual frames to
it. For instance, we will here capture 30 frames from the default webcam and write
them to the video file. 

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
my_stream <- stream(index = 0)
```

```{r, eval=FALSE}
for (i in seq_len(30)) {
  writeFrame(my_writer, readNext(my_stream))
}
```

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
release(my_stream)
```

`VideoWriter` objects can be released from memory as follows: 

```{r, eval=FALSE}
release(my_writer)
```

Note that `VideoWriter` objects are wrappers around [`OpenCV`](http://opencv.org/)
[`VideoWriter`](https://docs.opencv.org/3.4.1/dd/d9e/classcv_1_1VideoWriter.html)
objects. They will not persist between [`R`](https://cran.r-project.org) sessions.

### 2.2.2 - Images

Writing an image to a file is straightforward using the `write.Image` function.

[`Rvision`](https://github.com/swarm-lab/Rvision) will guess the format of the 
image file from the file extension. For instance, we will here capture 1 frame 
from the default webcam and write it to a temporary png image file. 

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
my_stream <- stream(index = 0)
```

```{r, eval=FALSE}
path_to_image <- paste0(tempfile(), ".png")
write.Image(readNext(my_stream), path_to_image)
```

```{r message=FALSE, warning=FALSE, include=FALSE, eval=FALSE}
release(my_stream)
```



