---
title: "3 - Basic operations"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{3 - Basic operations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 3.1 - Object properties

|     | `Video` | `Stream` | `Image` | `VideoWriter` |
|:--- |:------- |:-------- |:------- |:------------- |
| **Testing for** | `isVideo(x)` | `isStream(x)` | `isImage(x)` | `isVideoWriter(x)` |
| **Dimensions** | <code>[dim(x)](../reference/dim.Rcpp_Video.html)</code> | <code>[dim(x)](../reference/dim.Rcpp_Stream.html)</code> | <code>[dim(x)](../reference/dim.Rcpp_Image.html)</code> | - |
| **Number of rows** | <code>[nrow(x)](../reference/video_dimensions.html)</code> | <code>[nrow(x)](../reference/stream_dimensions.html)</code> | <code>[nrow(x)](../reference/nrow.Rcpp_Image.html)</code> | - |
| **Number of columns** | <code>[ncol(x)](../reference/video_dimensions.html)</code> | <code>[ncol(x)](../reference/stream_dimensions.html)</code> | <code>[ncol(x)](../reference/nrow.Rcpp_Image.html)</code> | - |
| **Number of channels** | - | - | `nchan(x)` | - |
| **Number of frames** | <code>[nframes(x)](../reference/video_dimensions.html)</code> | - | - | - |
| **Framerate** | `fps(x)` | - | - | - |
| **Codec** | `codec(x)` | - | - | - |
| **Bit depth** | - | - | `bitdepth(x)` | - |
| **Color space** | - | - | `colorspace(x)` | - |

---

## 3.2 - Displaying images

### 3.2.1 - Method compatible with base R

There are two methods for displaying images. The first method (`plot()`) offers 
full compatibility with base `R` plotting methods, which allows for drawing 
non-destructively over an image using `graphics::lines()`, `graphics::points()`, 
etc. For instance: 

```{r eval=FALSE, echo=TRUE}
# Find the path to the bunny.png image provided with Rvision
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")

# Load the image in memory
my_image <- image(filename = path_to_image)

# Plot the image using the base R plotting device
plot(my_image)

# Draw lines over the plotted image
abline(h = 400, col = "red", lwd = 5)
abline(h = 550, col = "red", lwd = 5)
abline(v = 430, col = "red", lwd = 5)
abline(v = 570, col = "red", lwd = 5)
```

Note that this method is rather slow and therefore not suitable for applications 
requiring quick refresh rates between successive images (e.g. when playing a 
video). 

### 3.2.2 - Method incompatible with base R

The second method (`display()`) uses [`OpenCV`](http://opencv.org/)'s capabilities 
to display images at high frame rates, making it suitable to play videos. For 
instance: 

```{r eval=FALSE, echo=TRUE}
# Find the path to the Balloon.mp4 video provided with Rvision
path_to_video <- system.file("sample_vid", "Balloon.mp4", package = "Rvision")

# Open the video file stream
my_video <- video(filename = path_to_video)

# Create an empty display window with width and height half of that of the video 
# (note: the display can hide behind other windows, especially on Mac)
newDisplay("My display", nrow(my_video) / 2, ncol(my_video) / 2) 

# Display images (note: it creates the display window if it doesn't exist yet)
# The 3rd argument ('delay') is the minimum time in ms during which the image 
# should be displayed before it can be replaced
for (i in 1:25) {
  display(readNext(my_video), "My display", 25, nrow(my_video) / 2, ncol(my_video) / 2)
}

# Close the display window
destroyDisplay("My display")
```

Note that the `display()` method is incompatible with base `R` plotting methods. 
However, you can use the drawing functions provided with `Rvision` to draw over 
images - destructively this time - before displaying them.

---

## 3.3 - Accessing parts of an image

You can access the pixel values of an image in exactly the same way you would 
access the values of a `base::array`. Actually, `Rvision` converts the image to 
a `base::array` before returning the desired values. For instance: 

```{r eval=FALSE, echo=TRUE}
my_image[]                  # Entire image
my_image[1, 1,]             # Bottom-left pixel
my_image[1,,]               # Bottom row of pixels
my_image[, 1,]              # Leftmost column of pixels
my_image[1:5, 1:5,]         # All pixels between the 1st and 5th row and column
my_image[c(TRUE, FALSE), c(TRUE, FALSE),]  # Every other row and column of pixels
```

The `pget()` function can be used to quickly access the values of specific pixels
by their x/y coordinates in the image. For instance:

```{r eval=FALSE, echo=TRUE}
# Return the pixel values at x/y coordinates (110,190) and (200,100)
pget(my_image, x = c(110, 200), y = c(190, 100))
```

This function returns a matrix where each column corresponds to one of the 
desired pixels and each row to the value of the pixels in each of the channels 
of the image (e.g. Blue, Red, Green).

---

## 3.4 - Replacing parts of an image

Images are not exactly like traditional `R` matrices. In particular, color 
images are made of multiple layers (also called channels), each corresponding to 
a dimension in a color space (e.g., the BGRA color space is composed of 4 
channels: Blue, Green, Red, and Alpha/Transparency). Therefore, when replacing a
pixel in a color image, it is necessary to provide values for each of these 
channels at once. This does not apply to grayscale images which have a single 
channel only and therefore behave similarly to regular monolayer matrices in 
base `R`. 

### 3.4.1 - Grayscale images

The traditional `<-` assignment operator can be used to replace parts of a 
grayscale image, like it would be used to replace any parts of a traditional 
`base::matrix`. For instance:

```{r eval=FALSE, echo=TRUE}
# Make a grayscale copy of the my_image object created earlier
my_gray_image <- changeColorSpace(my_image, "GRAY")

# Turn the bottom-left corner of the image black (0 in the grayscale space)
my_gray_image[1:250, 1:250] <- 0  
plot(my_gray_image)
```

Finally, the `pset()` function can be used to quickly set the values of specific 
pixels by their x/y coordinates in the image. For instance:

```{r eval=FALSE, echo=TRUE}
# Set a few pixels to white inside the black square added earlier
pset(my_gray_image, 50:100, 50:100, "white")
plot(my_gray_image)
```

### 3.4.1 - Color images

The `<-` assignment operator can also be used to replace parts of a color (i.e. 
multi-channel) image. However, instead of taking in single values for each pixel
to be replaced, it will take one value per pixel and per channel of the image. 
For instance: 

```{r eval=FALSE, echo=TRUE}
# Turn the bottom-left corner of the my_image object created earlier red. This 
# is a BGRA image, so it has 4 channels which new values need to be specified.
my_image[1:250, 1:250] <- c(0, 0, 255, 255) 
plot(my_image)
```

Note that if values are not specified for each channel, then the provided values
will be recycled. For instance:

```{r eval=FALSE, echo=TRUE}
# Turn the bottom-left corner of the my_image object completely transparent. 
# Transparency will appear as white when the image is plotted. 
my_image[1:250, 1:250] <- 0
plot(my_image)
```

The utility function `col2bgr()` provided by `Rvision` (not to be confused with 
`grDevices::col2rgb` on which `col2bgr()` is based) will conveniently convert any 
of the `R`-compatible color specifications to color codes compatible with 
`Rvision` images. For instance: 

```{r eval=FALSE, echo=TRUE}
# Turn the bottom-left corner of the my_image object blue. 
my_image[1:250, 1:250] <- col2bgr("blue", alpha = TRUE)
plot(my_image)
```

Finally, the `pset()` function can be used to quickly set the values of specific 
pixels by their x/y coordinates in the image. For instance:

```{r eval=FALSE, echo=TRUE}
# Set a few pixels to green inside the black square added earlier
pset(my_image, 50:100, 50:100, "green")
plot(my_image)
```
