---
title: "3 - Basic operations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r message=FALSE, warning=FALSE, include=FALSE}
library(Rvision)
```

__This page describes all basic operations on videos and images.__

---

+ [3.1 - Object properties ](#object-properties)
+ [3.2 - Displaying images](#displaying-images)
    + [3.2.1 - Method compatible with base R](#method-compatible-with-base-R)
    + [3.2.2 - Method incompatible with base R](#method-incompatible-with-base-R)
+ [3.3 - Accessing/replacing parts of an image](#access-replace)

---

## 3.1 - Object properties

|     | `Video` | `Stream` | `Image` | `VideoWriter` |
|:--- |:------- |:-------- |:------- |:------------- |
| **Testing for** | `isVideo(x)` | `isStream(x)` | `isImage(x)` | `isVideoWriter(x)` |
| **Dimensions** | <code>[dim(x)](../reference/dim.Rcpp_Video.html)</code> | <code>[dim(x)](../reference/dim.Rcpp_Stream.html)</code> | <code>[dim(x)](../reference/dim.Rcpp_Image.html)</code> | - |
| **Number of rows** | <code>[nrow(x)](../reference/video_dimensions.html)</code> | <code>[nrow(x)](../reference/stream_dimensions.html)</code> | <code>[nrow(x)](../reference/nrow.Rcpp_Image.html)</code> | - |
| **Number of columns** | <code>[ncol(x)](../reference/video_dimensions.html)</code> | <code>[ncol(x)](../reference/stream_dimensions.html)</code> | <code>[ncol(x)](../reference/nrow.Rcpp_Image.html)</code> | - |
| **Number of channels** | - | - | `nchan(x)` | - |
| **Number of frames** | <code>[nframes(x)](../reference/video_dimensions.html)</code> | - | - | - |
| **Framerate** | `fps(x)` | - | - | - |
| **Codec** | `codec(x)` | - | - | - |
| **Bit depth** | - | - | `bitdepth(x)` | - |
| **Color space** | - | - | `colorspace(x)` | - |

---

## 3.2 - Displaying images

There are two methods for displaying images. The first method offers full 
compatibility with base [`R`](https://cran.r-project.org) plotting methods, which 
allows for drawing non-destructively over an image using `graphics::lines()`, 
`garaphics::points()`, etc. This method is however is however rather slow and 
therefore not suitable for applications requiring quick refresh rates between 
successive images (e.g. when playing a video). 

The second method uses [`OpenCV`](http://opencv.org/)'s capabilities to display 
images at high frame rates, making it suitable to play videos for instance. It 
is however incompatible with base [`R`](https://cran.r-project.org) plotting 
functions. However you can use the [drawing functions](../drawing.html) provided 
with [`Rvision`](https://github.com/swarm-lab/Rvision) to draw over images 
- destructively this time - before displaying them.

### 3.2.1 - Method compatible with base R

```{r, message=FALSE}
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")
my_image <- image(filename = path_to_image)

plot(my_image)
abline(h = 400, col = "red", lwd = 5)
abline(h = 550, col = "red", lwd = 5)
abline(v = 430, col = "red", lwd = 5)
abline(v = 570, col = "red", lwd = 5)
```

### 3.2.2 - Method incompatible with base R

```{r, message=FALSE, eval=FALSE}
path_to_video <- system.file("sample_vid", "Balloon.mp4", package = "Rvision")
my_video <- video(filename = path_to_video)

# Create an empty display window (note: it can hide behind other windows)
newDisplay("My display", nrow(my_video) / 2, ncol(my_video) / 2) 

# Display images (note: it creates the display window if it doesn't exist yet)
# The 3rd argument ('delay') is the minimum time in ms during which the image 
# should be displayed before it can be replaced
for (i in 1:25) {
  display(readNext(my_video), "My display", 25, nrow(my_video) / 2, ncol(my_video) / 2)
}

# Close display
destroyDisplay("My display")

# Close all opened displays
destroyAllDisplays()
```

---

## 3.3 - Accessing/replacing parts of an image

You can access parts of an image in the same way you would do it with matrices 
and arrays in base [`R`](https://cran.r-project.org). 

```{r, message=FALSE, results=FALSE}
my_image[]          # Entire image
my_image[1, 1]      # Bottom-right pixel
my_image[1, ]       # Bottom row of pixels
my_image[, 1]       # Leftmost column of pixels
my_image[1:5, 1:5]  # All pixels between the 1st and 5th row and column
my_image[c(TRUE, FALSE), c(TRUE, FALSE)]  # Every other row and column of pixels
```

Note that images are not exactly like traditional matrices. Color images have 
multiple layers, each corresponsing to a dimension in the BGR or BGRA space 
(Blue, Green, Red, Alpha transparency). Therefore, when accessing a pixel of a
color image, you will obtain 3 or 4 values in return. Grayscale images, on the 
other hand, function similarly to regular monolayer matrices in base 
[`R`](https://cran.r-project.org). 

To replace parts of an image, you can use the `<-` assignment operator. With 
grayscale images, it works exactly like in base [`R`](https://cran.r-project.org) 
with matrices. 

```{r, message=FALSE, results=FALSE}
my_gray_image <- changeColorSpace(my_image, "GRAY")
my_gray_image[1:250, 1:250] <- 0  # Turn the corresponding pixels to black
plot(my_gray_image)
```

With color images however, you will need to supply 3 (for BGR images) or 4 (for 
BGRA images) values (or any multiple of 3 or 4 respectively). 

```{r, message=FALSE}
my_image[1:250, 1:250] <- c(0, 0, 255, 255) # Turn the corresponding pixels to red
plot(my_image)
```

The utility function `col2bgr` provided by [`Rvision`](https://github.com/swarm-lab/Rvision)
(not to be confused with `col2rgb` provided by the `grDevices` package and on 
which `col2bgr` is based) will conveniently convert any of the 
[`R`](https://cran.r-project.org)-compatible color specifications to color codes
compatible with [`Rvision`](https://github.com/swarm-lab/Rvision) images. 

```{r, message=FALSE}
my_image[1:250, 1:250] <- col2bgr("blue", alpha = TRUE) # Turn the corresponding pixels to blue
plot(my_image)
```
