---
title: "4 - In-place, copy, and target operations"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{4 - In-place, copy, and target operations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 4.1 - Better performances with in-place operations

While other image processing libraries for `R` are fast enough for most 
operations on single images and often offer more options than `Rvision` (but we
are working on closing the gap on that front), `Rvision` really shines in the 
high-speed processing of large volumes of images (e.g. video and camera streams).

Part of the reason is that [`OpenCV`](https://opencv.org/) - the computer vision
library on top of which `Rvision` is built - is designed and optimized for 
in-place operations, that is operations that change directly the content of an 
object in memory without making a copy of it. Indeed, making copies of objects 
in memory - especially of large ones like images - can add a significant amount
of time to an operation. For that reason, most of `Rvision`'s functions offer 
the option to perform their operations in-place, greatly increasing the 
processing speed of large volumes of images.

All operations that support in-place operations will have an argument named 
"target" that can be set to the value "self". For instance:

```{r eval=FALSE, echo=TRUE}
# Find the path to the bunny.png image provided with Rvision
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")

# Load the image in memory
my_image <- image(filename = path_to_image)

# Plot the original image
plot(my_image)

# Apply a Gaussian blur in place. Note that the operation will not return 
# anything and the result will replace the content of my_image.
gaussianBlur(my_image, 11, 11, 3, 3, target = "self")

# Plot original image after modification
plot(my_image)
```

---

## 4.2 - Preserving originals with copy operations

The downside of in-place operations is that the modification is destructive: the 
original object in memory is replaced bit for bit by its modified version (we 
are just talking about the RAM representation of the object here; the original 
file on your drive is never modified directly). 

However, it may be desirable in some cases to keep the original object in memory
untouched. For this purpose, all the functions in `Rvision` that can modify 
images also give users the option to perform their operations on a copy of the 
original image. This will be slower but resemble the usual behavior of `R` 
functions.

By default, all operations are copy operations but this can be explicitely 
specified by setting the "target" argument to "new". For instance:

```{r eval=FALSE, echo=TRUE}
# Find the path to the bunny.png image provided with Rvision
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")

# Load the image in memory
my_image <- image(filename = path_to_image)

# Apply a Gaussian blur to a copy of the image. 
my_blurred_copy <- gaussianBlur(my_image, 11, 11, 3, 3, target = "new")

# Plot original image
plot(my_image)

# Plot blurred copy 
plot(my_blurred_copy)
```

---

## 4.3 - The best of both worlds: target operations

But what if I want to preserve the original object AND be fast at the same time? 

Then you are in luck. `Rvision` - still thanks to `OpenCV` - also offers "target" 
operations, that is operations which results can be stored in another image in
memory (the target) as long as that image was pre-allocated before the operation
takes place and that it has the same memory footprint as the original image. 
This is as fast as in-place operations and it does not modify the original image;
instead, it modifies a separate image that serves as a placeholder for the 
results. 
 
All operations that support target operations will have an argument named 
"target" that can be set to the name of the target image. For instance:

```{r eval=FALSE, echo=TRUE}
# Find the path to the bunny.png image provided with Rvision
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")

# Load the image in memory
my_image <- image(filename = path_to_image)

# Create a zero-filled target image with the memory footprint of the original
my_target <- zeros(nrow(my_image), ncol(my_image), nchan(my_image), bitdepth(my_image))

# Apply a Gaussian blur to the original image but store the results in the 
# target. Note that the operation will not return anything and the result will 
# replace the content of my_target.
gaussianBlur(my_image, 11, 11, 3, 3, target = my_target)

# Plot original image
plot(my_image)

# Plot target
plot(my_target)
```

---

## 4.4 - When to use in-place, copy, and target operations? 

If you are going to process a single image (or a small batch of images) with 
only one or a few simple operations, then copy operations are probably the best
choice. They will be fast enough (a fraction of a second at most) and more 
familiar to `R` users since they essentially mimic the behavior of traditional 
`R` functions by always returning a new object. They are also very useful when 
troubleshooting your pipeline since every step will be saved in a separate 
object and can therefore be inspected individually for problems. 

If you are going to process a large number of images (typically, when processing
a video or camera stream) and you only care about preserving the final result of 
the pipeline, then in-place operations will be the best choice. They will 
provide the best performance both in terms of speed and overall memory footprint. 

Finally, if you are going to process a large number of images but need to 
preserve the original image or the result of an intermediary step in your 
pipeline (e.g. in order to reuse it later in your pipeline), then target 
operations should be preferred. 

Note that the different operation types are not mutually exclusive and can be 
used together in the same processing pipeline for increased flexibility.
