---
title: "5 - GPU operations"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{5 - GPU operations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 5.1 - Stupid-fast performances with GPU operations

Since version `0.6.0`, running operations on the GPU (Graphics Processing Units) 
of a computer is possible and - we want to believe - very easy to implement. 
GPUs are designed to allow for parallel pixel processing. This can result in 
significant gains in performance for operations that can benefit from massive 
parallelization, making them in some cases several orders of magnitude faster 
than the same operation on the CPU (Central Processing Unit) of the computer. 

[`OpenCV`](https://opencv.org/) - the computer vision library on top of which 
`Rvision` is built - can handle operations on the GPU using two popular 
frameworks: [`OpenCL`](https://www.khronos.org/opencl/) and 
[Nvidia's `CUDA`](https://developer.nvidia.com/cuda-zone). `OpenCL` being open, 
royalty-free, and cross-platform, it is available to more users' machines (`CUDA` 
is proprietary and restricted to `Nvidia` graphics cards). We have, therefore,
chosen to use that framework in `Rvision`. This is not to say that we will never 
consider adding `CUDA` support to `Rvision`; we probably will - time permitting 
- because it is reportedly faster than `OpenCL` in many instances. We just 
decided to start with the framework that is available to more users immediately. 

---

## 5.2 - Enabling GPU operations in `Rvision`

By default (and for good reasons too), `Rvision` loads images it receives in the 
memory of the CPU. Before using GPU-accelerated functions, the image must first 
be copied to the GPU memory. This can be done very easily as follows:

```{r eval=FALSE, echo=TRUE}
# Find the path to the bunny.png image provided with Rvision
path_to_image <- system.file("sample_img", "bunny.png", package = "Rvision")

# Load the image in memory
my_image <- image(filename = path_to_image)

# Copy the image to GPU memory
my_image$toGPU()
```

Once this is done, `Rvision` (`OpenCV` really) will automatically use the 
GPU-accelerated version of each operation on the image if it is available. 
Otherwise, it will default to using the CPU version. That is it, there is 
nothing more complicated to do to take advantage of the processing speed 
of the GPU. 

In the case where a function accepts multiple images as arguments (e.g. when 
using a target image to save the result of an operation on the original image), 
then it does not matter in most cases if all the images are on the CPU or on the 
GPU (in the rare cases when it does matter, an error will be thrown and you can 
adjust your pipeline accordingly). However, better performances will probably be
achieved if all the images use the memory of the same processor. 

Finally, if you need to copy back the image to the CPU memory (e.g. if you want
to then transfer the image to a `base::array`), you can simply apply the reverse 
operation as follows: 

```{r eval=FALSE, echo=TRUE}
# Copy the image back to CPU memory
my_image$fromGPU()
```

---

## 5.3 - Caveats to GPU operations

Nothing comes for free and there are a few caveats to using GPU-accelerated 
operations. 

First, it will only work if your computer has a GPU (even if a basic one 
integrated with the CPU) and that the `OpenCL` drivers for the GPU are installed.
If `OpenCL` is not available on your system, `Rvision` will throw an error when 
you use the `$toGPU` function. `OpenCL` should be available on most traditional 
computers (including laptops) with most operating systems. Check the 
documentation of your computer and operating system to figure out whether `OpenCL` 
is available/installed on your system and how to install it if necessary (in 
which case, you will probably need to recompile `OpenCV` using the 
`ROpenCVLite::installOpenCV` function). `OpenCL` will probably not be available
on shared servers that do not provide GPU access to their users. Check with your 
server administrator if that it the case. 

While GPU-accelerated operations can be much faster than their CPU-based 
equivalent, not all operations can be efficiently run on a GPU. Some operations
provided by `Rvision`/`OpenCV` will actually be slower on the GPU than the CPU.
Moreover, performances will highly depend on the abilities of your GPU. A basic,
integrated GPU will perform a lot slower than a dedicated graphics card for 
instance, and in many cases not faster than the CPU. Test your pipeline 
carefully to decide whether it is worth running all or parts of it on the GPU, 
or to keep everything on the CPU instead. 

Copying an image to/from the GPU comes with a time penalty. It is best if your 
pipeline avoids performing too many copying operations. Ideally, the GPU/CPU
copies should all be created/pre-allocated before starting the pipeline for 
better performances. 

Finally, GPU operations will be slow the first time they are run during a 
session. This is because `OpenCV` needs to compile the corresponding functions 
for the specific graphics device it will use to run the operations. Therefore, 
GPU operations should be avoided if they will only be used a handful of times 
during a session. They are better suited for heavily repeated operations, on 
large volumes of images where their speed gains can quickly compensate for the 
time penalty they incur at the start of the pipeline. Note that a possible way 
to mitigate this problem is to do a "warm-up" run of all the functions at the 
start of a session, before the pipeline starts. 
