---
title: "6 - Capture queues"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{6 - Capture queues}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 6.1 - Increasing performances with capture queues

Retrieving a single frame from a video or camera stream is fast, but it is not 
instantaneous because `Rvision` (and therefore `OpenCV` in the background) needs
to grab and decode each frame before it can be used for further processing. For 
most applications, that small time penalty is not an issue; losing a few 
milliseconds per frame will not be felt by a user processing short videos, low 
resolution videos for instance. 

However, the frame decoding time will increase significantly with the resolution 
of the video/camera stream; and for applications requiring the processing of 
long videos or live camera feeds, the total time lost retrieving frames will 
quickly increase. Moreover, the process of retrieving frames is blocking: while
`Rvision` grabs and decodes a frame, it cannot work on other frames that were 
previously captured. 

A solution to this issue is to make use of the multi-tasking ability of most 
modern computers to perform the frame retrieving process in parallel with the 
rest of the image processing. The principle is fairly simple: one processing 
thread (thread 1) is in charge of grabbing, decoding, and storing frames in a 
shared dynamic queue (or buffer); a second thread (thread 2) is in charge of 
processing these pre-loaded frames further; while thread 2 is working on a frame, 
thread 1 can keep filling up the queue with new frames so that thread 2 never 
has to stop and wait for a new frame to be retrieved. On a multi-core/processor 
computer (most computers nowadays), thread 1 and 2 can be operated in parallel 
on separate cores/processors, effectively reducing the frame waiting time to 
near-zero for thread 2 (but see caveats section below). 

Unfortunately, `R` is a single-threaded language, meaning that it cannot 
natively create and operate multiple parallel processing threads. Fortunately,
the [`C++` language](https://www.cplusplus.com/) that is used in the background
by `Rvision` can do it. We used that ability - and some precautions to avoid 
memory access conflicts with `R` - to create the `Queue` class of objects which 
job is to pre-load in memory video and camera stream frames without blocking the 
execution of the main (and unique) `R` processing thread, and then give the 
main `R` thread near-instantaneous access to these pre-loaded frames when 
required. Once a frame has been collected by the main `R` thread, it is removed
from the queue to make space for another frame. 

--- 

## 6.2 - Creating a queue

`Queue` objects are created using the `queue()` function. For instance:  

```{r eval=FALSE, echo=TRUE}
# Find the path to the Balloon.mp4 video provided with Rvision
path_to_video <- system.file("sample_vid", "Balloon.mp4", package = "Rvision")

# Open the video file stream
my_video <- video(filename = path_to_video)

# Create a queue of frames 
my_buf <- queue(my_video, size = 10, delay = 1000, overflow = "pause")
```

The `queue()` function can take 4 parameters:

+ `x` corresponds to the source of the frames to be queued. It should either be 
a `Video` or a `Stream` object. 
+ `size` corresponds to the number of frames that the queue can store at any one
time. By default, a `Queue` object will be able to hold 10 frames at once. If 
you increase the value of the `size` parameter, the `Queue` object will use up 
more RAM as a result, so be mindful of your computer's resources. 
+ `delay` corresponds to the time (in microseconds) between two queue update 
cycles. During an update cycle, the `Queue` object checks whether it is full or 
not and whether frames have been collected by the main `R` thread. If it is not 
full, it retrieves a new frame from the source and stores it; if frames have 
been collected by the main `R` thread, it removes them from the queue; it then 
waits the duration set by `delay` before starting a new update cycle. Reducing 
the value of `delay` will increase the frequency of the update cycles but will 
also increase the computational load of the core/processor running the queuing 
thread.
+ `overflow` corresponds to the behavior the `Queue` object should adopt once it
is full. By default, the queuing process will "pause" (that is stop retrieving 
new frames from the source and storing them) until a frame is collected by the 
main `R` thread. This is the behavior of choice for video processing as it does 
not skip any frame while ensuring that the queue storage memory does not grow 
more than what the user wants. The queuing process can also "replace" the oldest
frame in the queue with a new one. This is usually a good choice for processing
live camera stream as it ensures that the queue storage memory does not grow 
more than what the user wants. However, frames may be skipped if the main `R` 
thread cannot collect and process the frames faster than they can be retrieved
and stored by the queuing thread. Finally, the queuing process can "grow" the 
queue by doubling its size each time it fills up. This behavior allows for not 
pausing the retrieval process and avoids frame skipping but it is very much NOT
recommended unless you know what you are doing. Indeed, this can lead to 
excessive RAM usage and decreased computing performance across the board. 

Once a `Queue` object is created, it starts immediately filling up with images 
retrieved from the `Video` or a `Stream` source object. Note that, if you had 
previously read frames from a `Video` source object before creating the `Queue` 
object, the latter will start retrieving frames from where you left off (e.g. if
you have already read the first 10 frames, the queue will start filling up from 
the 11th). 

Once a `Queue` object is not required anymore, it can be released from memory as 
follows: 

```{r eval=FALSE, echo=TRUE}
release(my_buf)
```

---

## 6.3 - Using a `Queue` object

The main purpose of a `Queue` object is to pre-load and store frames for fast 
access later on. The pre-loading and storing happens by itself in the background
so you do not need to take care of this. Collecting a frame from the queue into 
the main `R` thread can be done using the `readNext()` function. For example:

```{r eval=FALSE, echo=TRUE}
# Collect the next available frame from the queue and store it in a new 
# Image object
frame <- readNext(my_buf)

# Collect the next available frame from the queue and store it in an existing 
# Image object
readNext(my_buf, target = frame)
```

At any time, you can check the state of the queue as follows: 

```{r eval=FALSE, echo=TRUE}
# Is the queue empty?
empty(my_buf)

# Is the queue full?
full(my_buf)

# What is the current number of frames in the queue? 
length(my_buf)

# What is the maximum number of frames that the queue can hold?
capacity(my_buf)

# What is the index of the next frame available? (for video queues only)
frame(my_buf)

# What are the dimensions of the queue? 
dim(my_buf)
nrow(my_buf)
ncol(my_buf)
```

---

## 6.4 - Caveats to using `Queue` objects

`Queue` objects can be very useful to speed up the processing of camera streams
and long videos, especially if their resolution is high (HD or above). For short
videos or low resolution videos and camera streams, the benefits of using a 
`Queue` object will be very limited because retrieving frames from such sources 
is already very fast. 

If the processing of the frames on the main `R` thread is much faster than the 
retrieving and storing of the frames inside a `Queue` object, you risk emptying 
the queue faster than in can fill up. If that is the case, `readNext()` will 
display a warning to this effect but will not return an error (as it would, for
instance, when reaching the end of a video file). In addition, it will not 
create a new image or modify the target image. You will need to make sure that 
your code can catch these warnings and act accordingly (e.g., wait a turn before
trying again). You can reduce the risks of the queue becoming empty by (1)
increasing the size of the queue when creating it, (2) making sure the queue has 
completely filled up before starting the processing of the frames, and (3) 
reducing the delay between to queue updates. In any case, `Queue` objects are 
better suited for and more beneficial in cases where the processing of the 
frames is as slow or slower than the frame retrieving process. 

Finally, `Queue` objects read frames from existing `Video` and `Stream` objects. 
This means that they modify the state of these source objects the same way a 
user would by reading directly from them. This has a few of consequences that 
require some attention when using `Queue` object. First, a `Queue` object will 
start reading the source object from whichever state you have left it off. For 
instance, if you have read the first 10 frames of a video before passing it to a
`Queue` object, the `Queue` object will start reading it from the 11th frame. 
Conversely, if you pass a newly created `Video` object to a `Queue` object with 
a size of 10, the `Queue` object will immediately start reading and storing the 
first 10 frames. If you then decide to read a frame directly from the `Video` 
object, it will return its 11th frame, not its first one. Therefore, to avoid 
competing with a `Queue` object on a given `Video` object, it is strongly 
recommended to avoid creating code that mix reading frames from a `Queue` object
and reading frames from the `Video` object that `Queue` object is using.
