---
title: "7 - Video stacks"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{7 - Video stacks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## 7.1 - Introducting video stacks

Starting with `Rvision` version 0.7.0, users can create video stacks. A video 
stack (`VideoStack`) is essentially a list of videos with similar properties (in 
particular, the same dimensions) that behaves almost entirely like a single 
continuous `Video` object. For that reason, `VideoStack` objects are 
particularly useful for applications requiring the combination of multiple video
files and/or `Video` objects. For instance, video stacks can greatly simplify 
the merging of two separate video files into a single one, or the processing of 
a continuous sequence that was split into multiple files during recording.

In this vignette, we will briefly describe how to create and manipulate 
`VideoStack` objects. Since their operation is very similar to that of `Video` 
objects, using `VideoStack` objects should not require much additional learning
for a regular user of `Rvision`.

---

## 7.2 - Creating video stacks

You can create a video stack by passing to the `videoStack` function a 
combination of existing `Video` objects and/or character strings pointing toward 
the location of video files on your computer. For instance: 

```{r eval=FALSE, echo=TRUE}
# Find the path to the Balloon.mp4 video provided with Rvision
path_to_video <- system.file("sample_vid/Balloon.mp4", package = "Rvision")

# Open the video file stream
my_video <- video(filename = path_to_video)

# Create a video stack from a character string and a Video object
my_video_stack <- videoStack(path_to_video, my_video)
```

The paths to video files and the existing `Video` objects can be passed as 
separate arguments, or as a list or a vector. If a path is provided, the 
corresponding video will be opened before being added to the stack. If a `Video`
object is provided, a pointer to the original `Video` object will be added to 
the stack (i.e., they will share the same reader head and any modification to 
the position of the reader head in the stack will be applied to the original 
object, and vice versa). 

A video stack can contain as many videos as desired, with the only restriction
that they should all have the same dimensions. 

Once a `VideoStack` object is not required anymore, it can be released from 
memory as follows: 

```{r eval=FALSE, echo=TRUE}
release(my_video_stack)
```

---

## 7.3 - Using video stacks

Once a `VideoStack` object is created, it can be used as a `Video` object with 
the same number of frames as the sum of the frames in the `Video` objects that 
it contains. For instance, to read a specific frame of the stack, you can use 
the `readNext()` and `readFrame()` functions, as you would normally do with 
`Video` objects: 

```{r eval=FALSE, echo=TRUE}
# Capture the next available frame from the my_video_stack object created earlier
my_image <- readNext(my_video_stack)

# Capture frame 100 from the my_video_stack object created earlier
my_image <- readFrame(my_video_stack, 100)
```

At any time, you can check the state of the video stack as follows: 

```{r eval=FALSE, echo=TRUE}
# What is the current number of videos in the stack? 
length(my_video_stack)

# What is the number of frames in the stack? 
nframes(my_video_stack)

# What is the index of the next frame available? 
frame(my_video_stack)

# What are the dimensions of the queue? 
dim(my_video_stack)
nrow(my_video_stack)
ncol(my_video_stack)

# What are the frame rates of the videos in the stack?
fps(my_video_stack)

# What are the codecs of the videos in the stack?
codec(my_video_stack)
```

Finally, you can access individually, add, or replace videos in the stack as 
follows:

```{r eval=FALSE, echo=TRUE}
# Access the first video in the stack individually
my_video_stack[[1]]

# Replace the first video in the stack 
my_video_stack[[1]] <- my_video

# Add a video at the end of the stack 
my_video_stack[[length(my_video_stack) + 1]] <- my_video

# Remove a video from the stack
my_video_stack[[2]] <- NULL
```
